Dirichlet discretization of Neumann problems

### 2020-01-07 paper_lap/05apparatus.tex
updating manuscript

The following lemma relates the discretization of the inverse of an operator to the adjoint of the discretization of its inverse. Its proof follows directly from the definition of the adjoint and is omitted.
\begin{lemma}\label{lem_adjm}
Suppose $A: L^2([0,L]) \to L^2([0,L])$ is a bounded invertible operator and that $A_\epsilon$ is an operator such that
\begin{align}
\left| \langle f,A^{-1} g \rangle - \langle f,A^{-1}_\epsilon g \rangle \right| \le \epsilon \|f\| \|g\|,
\end{align}
for all $f$ and $g$ in some subspace $S \subset L^2([0,L]).$ Here $\langle \cdot, \cdot \rangle$ denotes the inner product on $L^2([0,L])$ and $\| \cdot\|$ denotes the norm for $L^2([0,L]).$ Then, for all functions $f$ and $g$ in $S$
\begin{align}
\left| \langle f,(A^{-1})^* g \rangle - \langle f,\left(A^{-1}_\epsilon\right)^* g \rangle \right| \le \epsilon \|f\| \|g\|
\end{align}
where $\,{{}^*}\,$ denotes the adjoint.
\end{lemma}

@@

\begin{remark1}
We observe that if the solution to the boundary value problem is being calculated at a point $\by \in \mathbb{R}^2\setminus \Omega$ more than {\color{red}two panel lengths} away from the boundary curve $\Gamma$ then the Neumann density $\sigma$ obtained using the above result will give an accuracy of $\epsilon,$ ie. the function $K(\by,\gamma(t)) \in S_\epsilon.$ Thus accurate values of the solution in the far-field can be obtained almost immediately.
\end{remark1}
\begin{remark1}
Similarly, if the point $\by \in \mathbb{R}^2 \setminus \Omega$ at which the solution to the Neumann boundary value problem is to be calculated lies close to a smooth panel then the density $\sigma$ near that point can be interpolated to a finer set of quadrature points and the value of $u(\by)$ can once again be obtained to precision $\epsilon.$ We note, however, that in general the density in the vicinity of a corner cannot be interpolated accurately. This follows from the fact that the interpolation scheme constructed is only guaranteed to interpolate the powers arising in the Dirichlet problem accurately near the corner. The collection of singular powers arising in Neumann problems contain negative powers which are not contained in this set and hence are not interpolated accurately.
\end{remark1}

@@

. Additionally, we denote the discretization of the interior Dirichlet operator (using the custom quadratures described in Section \ref{sec:disc_dir}) by $A.$ Let $\underline{f} = (f_1,\dots,f_N)^T$ where $f_i = f(x_i) \sqrt{w_i}$ and $f: \partial \Omega \to \mathbb{R}$ is the right-hand side of the exterior Neumann problem. Finally, let $\underline{\sigma}$ be the approximation to the density (scaled by the square roots of the weights) obtained by solving the linear system
\begin{align}\label{eqn:adj_linalg}
A^T \underline{\sigma} = \underline{f}.
\end{align}
For notational convenience we let $\gamma:[-\delta,L-\delta] \to \partial \Omega$ be a counterclockwise arclength parametrization of $\partial \Omega$ such that  $\gamma(0)$ corresponds to a vertex and $\gamma[-\delta,\delta]$ corresponds to a corner panel.

For a panel $\gamma([s_1,s_2])$ with discretization nodes $x_i,\dots,x_{i+M}$ corresponding to a Gauss-Legendre panel the density is smooth and thus it is expected to be well-represented in the basis of Legendre polynomials (shifted and scaled to the interval $[s_1,s_2]$). Hence standard interpolation techniques can be used to obtain an accurate approximation to the density $\sigma$ on the interval $s_1 \le s\le s_2$ (see \cite{} for a detailed description of this process and quantitative estimates on the accuracy of the result). Typically we use 16th order Gauss-Legendre panels and choose their sizes so that their length is no more than their distance to the nearest corner. {\color{red}This latter choice guarantees that for any $\epsilon>0$ there exists an $M$ such that if the Gauss-Legendre panels are discretized using an $M$-point Gauss-Legendre rule then the density on that panel can be interpolated to relative precision $\epsilon$ in an $L^2$-sense.}

>>>>

The following lemma relates the discretization of the inverse of an operator to the adjoint of the discretization of its inverse. Its proof follows directly from the definition of the adjoint and is omitted.
\begin{lemma}\label{lem_adjm}
Suppose $A: L^2([0,L]) \to L^2([0,L])$ is a bounded invertible operator and that $A_\epsilon$ is an operator such that
\begin{align}
\left| \langle f,A^{-1} g \rangle - \langle f,A^{-1}_\epsilon g \rangle \right| \le \epsilon \|f\| \|g\|,
\end{align}
for all $f$ and $g$ in some subspace $S_{\epsilon} \subset L^2([0,L]).$ Here $\langle \cdot, \cdot \rangle$ denotes the inner product on $L^2([0,L])$ and $\| \cdot\|$ denotes the norm for $L^2([0,L]).$ Then, for all functions $f$ and $g$ in $S_{\epsilon}$
\begin{align}
\left| \langle f,(A^{-1})^* g \rangle - \langle f,\left(A^{-1}_\epsilon\right)^* g \rangle \right| \le \epsilon \|f\| \|g\|
\end{align}
where $\,{{}^*}\,$ denotes the adjoint.
\end{lemma}

@@

\begin{remark1}
We observe that if the solution to the boundary value problem is being calculated at a point $\by \in \mathbb{R}^2\setminus \Omega$ more than one panel length away from the boundary curve $\Gamma$ then the Neumann density $\sigma$ obtained using the above result will give an accuracy of $\epsilon,$ ie. the function $K(\by,\gamma(t)) \in S_\epsilon.$ Thus accurate values of the solution in the far-field can be obtained almost immediately.
\end{remark1}
\begin{remark1}
Similarly, if the point $\by \in \mathbb{R}^2 \setminus \Omega$ at which the solution to the Neumann boundary value problem is to be calculated lies close to a smooth panel then the density $\sigma$ near that point can be interpolated to a finer set of quadrature points and the value of $u(\by)$ can once again be obtained to precision $\epsilon.$ We note, however, that in general the density in the vicinity of a corner cannot be interpolated accurately. This follows from the fact that the interpolation scheme constructed is only guaranteed to interpolate the powers arising in the Dirichlet problem accurately near the corner. The collection of singular powers arising in Neumann problems contain negative powers which are not contained in this set and hence are not interpolated accurately.
\end{remark1}

@@

Additionally, we denote the discretization of the interior Dirichlet operator (using the custom quadratures described in Section \ref{sec:disc_dir}) by $A.$ Let $\underline{f} = (f_1,\dots,f_N)^T$ where $f_i = f(x_i) \sqrt{w_i}$ and $f: \partial \Omega \to \mathbb{R}$ is the right-hand side of the exterior Neumann problem. Finally, let $\underline{\sigma}$ be the approximation to the density (scaled by the square roots of the weights) obtained by solving the linear system
\begin{align}\label{eqn:adj_linalg}
A^T \underline{\sigma} = \underline{f}.
\end{align}
For notational convenience we let $\gamma:[-\delta,L-\delta] \to \partial \Omega$ be a counterclockwise arclength parametrization of $\partial \Omega$ such that  $\gamma(0)$ corresponds to a vertex and $\gamma[-\delta,\delta]$ corresponds to a corner panel.

For a panel $\gamma([s_1,s_2])$ with discretization nodes $x_i,\dots,x_{i+M}$ corresponding to a Gauss-Legendre panel the density is smooth and thus it is expected to be well-represented in the basis of Legendre polynomials (shifted and scaled to the interval $[s_1,s_2]$). Hence standard interpolation techniques can be used to obtain an accurate approximation to the density $\sigma$ on the interval $s_1 \le s\le s_2$ . Typically we use 16th order Gauss-Legendre panels and choose their sizes so that their length is no more than their distance to the nearest corner. This latter choice guarantees that for any $\epsilon>0$ there exists an $M$ such that if the Gauss-Legendre panels are discretized using an $M$-point Gauss-Legendre rule then the density on that panel can be interpolated to relative precision $\epsilon$ in an $L^2$-sense.
(We discuss a sketch of a proof in~\cref{sec:appb})

### 2020-01-07 paper_lap/manuscript.tex
updating manuscript

\input{01intro}
\input{02mprelim}
\input{03bvp}
\input{04nprelim}
\input{05apparatus}
\input{06num}
\input{07conc}
\section*{Appendix A}
Here we give explicit bounds for the rate of convergence of the contribution of the rest of the boundary to a corner panel. In particular, given a polygonal domain with boundary $\Gamma,$ let $\bx$ denote a vertex of $\Gamma$ and $C = \Gamma \cap B_r(\bx),$ where $B_r(\bx)$ is the ball of radius of $r$ centered at $\bx.$ We choose $r$ so that $\Gamma \cap B_{2r}(\bx)$ corresponds to a wedge with internal angle $\pi \alpha$ and side lengths $2r.$

@@

\begin{theorem}
Let $\Gamma$ be the boundary of a polygon and $\bx$ be a vertex. Let $r>0$ be a real number such that $\Gamma \cap B_{2r}(\bx)$ corresponds to a wedge with internal angle $\pi \alpha$ and side lengths $2r,$ where $B_R(\bx)$ denotes a ball of radius $R$ centered at $\bx.$ Let $L$ denote the length of $\Gamma$ and $\gamma: [-L/2,L/2]\to \Gamma$ be an arclength counterclockwise parameterization of $\Gamma$ such that $\gamma(0) = \bx.$ Finally, for any $f \in L^2(\Gamma),$ let $H:[0,r] \to \mathbb{R}$ be the function defined by
\begin{align}
H(t) = \int_{\Gamma\backslash B_{2r}(\bx)} K(\gamma(t),\gamma(s)) f(s)\,{\rm d}s.
\end{align}
Then $H$ is analytic in a neighborhood of $0$ with Taylor series coefficients $\{ a_n\}$ satisfying
\begin{align}
|a_n| \le \frac{\sqrt{L}}{2^nr^{n+1}} \|f\|_{L_2(\Gamma \backslash B_{2r}(\bx))}.
\end{align}
\end{theorem}
\begin{proof}
Without loss of generality we can assume that $\Gamma$ is shifted, oriented and parameterized so that $\bx = \gamma(0)= 0$ and the leg of the wedge corresponding to positive $t$ is oriented along the positive $x$ axis. Then
\begin{align}
H(t) = \int_{\Gamma\backslash B_{2r}(\bx)} K(\gamma(t),\gamma(s)) f(s)\,{\rm d}s = \int_{\Gamma\backslash B_{2r}(\bx)} \frac{y(s)}{(t-x(s))^2+y(s)^2}f(s)\,{\rm d}s.
\end{align}
Since $\|\gamma(s)-\gamma(t)\| >2r$ it follows that
\begin{align}
H(t) =\int _{\Gamma\backslash B_{2r}(\bx)}\sum {t^n i}\left(\frac{1}{(x(s)+i y(s))^{n+1}}-\frac{1}{(x(s)-i y(s))^{n+1}} \right)\,{\rm d}s.
\end{align}
In particular, $H$ has a Taylor series about $t=0,$
\begin{align}
\sum_{n=0}^\infty c_n \left(\frac{t}{2r}\right)^n
\end{align}
where
\begin{align}
|c_n| \le  \frac{\sqrt{L}}{r} \|f\|_{L_2(\Gamma \backslash B_{2r}(\bx))}.
\end{align}
\end{proof}

\end{document}

>>>>

\input{01intro}
\input{02mprelim}
\input{03bvp}
\input{04nprelim}
\input{05apparatus}
\input{06num}
\input{07conc}

\section*{Appendix A}
Here we give explicit bounds for the rate of convergence of the contribution of the rest of the boundary to a corner panel. In particular, given a polygonal domain with boundary $\Gamma,$ let $\bx$ denote a vertex of $\Gamma$ and $C = \Gamma \cap B_r(\bx),$ where $B_r(\bx)$ is the ball of radius of $r$ centered at $\bx.$ We choose $r$ so that $\Gamma \cap B_{2r}(\bx)$ corresponds to a wedge with internal angle $\pi \alpha$ and side lengths $2r.$

@@

\begin{theorem}
Let $\Gamma$ be the boundary of a polygon and $\bx$ be a vertex. Let $r>0$ be a real number such that $\Gamma \cap B_{2r}(\bx)$ corresponds to a wedge with internal angle $\pi \alpha$ and side lengths $2r,$ where $B_R(\bx)$ denotes a ball of radius $R$ centered at $\bx.$ Let $L$ denote the length of $\Gamma$ and $\gamma: [-L/2,L/2]\to \Gamma$ be an arclength counterclockwise parameterization of $\Gamma$ such that $\gamma(0) = \bx.$ Finally, for any $f \in L^2(\Gamma),$ let $H:[0,r] \to \mathbb{R}$ be the function defined by
\begin{align}
H(t) = \int_{\Gamma\backslash B_{2r}(\bx)} K(\gamma(t),\gamma(s)) f(s)\,{\rm d}s.
\end{align}
Then $H$ is analytic in a neighborhood of $0$ with Taylor series coefficients $\{ a_n\}$ satisfying
\begin{align}
|a_n| \le \frac{\sqrt{L}}{2^nr^{n+1}} \|f\|_{L_2(\Gamma \backslash B_{2r}(\bx))}.
\end{align}
\end{theorem}
\begin{proof}
Without loss of generality we can assume that $\Gamma$ is shifted, oriented and parameterized so that $\bx = \gamma(0)= 0$ and the leg of the wedge corresponding to positive $t$ is oriented along the positive $x$ axis. Then
\begin{align}
H(t) = \int_{\Gamma\backslash B_{2r}(\bx)} K(\gamma(t),\gamma(s)) f(s)\,{\rm d}s = \int_{\Gamma\backslash B_{2r}(\bx)} \frac{y(s)}{(t-x(s))^2+y(s)^2}f(s)\,{\rm d}s.
\end{align}
Since $\|\gamma(s)-\gamma(t)\| >2r$ it follows that
\begin{align}
H(t) =\int _{\Gamma\backslash B_{2r}(\bx)}\sum {t^n i}\left(\frac{1}{(x(s)+i y(s))^{n+1}}-\frac{1}{(x(s)-i y(s))^{n+1}} \right) f(s)\,{\rm d}s.
\end{align}
In particular, $H$ has a Taylor series about $t=0,$
\begin{align}
\sum_{n=0}^\infty c_n \left(\frac{t}{2r}\right)^n
\end{align}
where
\begin{align}
|c_n| \le  \frac{\sqrt{L}}{r} \|f\|_{L_2(\Gamma \backslash B_{2r}(\bx))}.
\end{align}
\end{proof}

\section{Appendix B}
\label{sec:appb}
In this section, we demonstrate that for a panel which is away from the corner, which is discretized using $M$ Gauss-Legendre nodes, where the solution is computed using the adjoint of a Dirichlet discretization, the density can be interpolated 
accurately at any point on the panel. As before let $\Omega$ denote a polygonal domain with boundary $\Gamma$. Let
$L$ denote the length of the boundary, and let $\gamma: [-L/2,L/2]$ denote an arc-length parameterization of the boundary.
Let $\gamma([s_{1},s_{2}])$ be a panel discretized using $M$ Gauss-Legendre nodes be such that if $r_{0}$ denotes
the distance to the closest corner on the particular edge, the length of the panel $L_{0}<r_{0}/2$, i.e. the panel
is separated from the nearest corner by at least two times its own length. Further assume that no point on the boundary $\Gamma$ which is not on the same edge as as $\gamma([s_{1},s_{2}])$ is contained in some $\rho-$Bernstein ellipse
corresponding to the line segment $\gamma([s_{1},s_{2}])$ with $\rho>1$. Furthermore suppose that all of the panels away from the corner and not on the edge containing the segment $\gamma([s_{1},s_{2}])$ are such that any point on $\gamma([s_{1},s_{2}])$ is not contained in their $\rho-$ Bernstein ellipse.
Both of these assumptions can always be met by choosing sufficiently small panels away from the corners. Finally 
assume that all panels at any corner $r_{c}$ are such that any point on $\gamma([s_{1},s_{2}])$ is at least $2r_{c}$ 
away from the corner vertex. Under all of these assumptions, it can be shown that the accuracy in computing the
Legendre coefficients of the density corresponding to the Neumann problem, is the same as the accuracy
of a related Dirichlet problem where the boundary data is resolved with controlled error on the given discretization.

Let $A$ denote the operator corresponding to interior Dirichlet problem using a double layer potential. Let $g$ denote the
right hand side for the Neumann problem, let $\sigma$ denote the corresponding solution. Let $f$ be a Legendre polynomial of
degree $n$ scaled to the panel $\gamma_{[s_{1},s_{2}]}$ and $0$ everywhere else. Then
\begin{equation}
\left<\sigma, f \right> = \left< (A^{T})^{-1} g, f\right> = \left< g, A^{-1} f \right> = <g, \sigma_{f}> \, ,
\end{equation}
where $\sigma_{f}$ is the solution of the interior Dirichlet problem with boundary data $f$ using a double layer potential.

Using~\cref{lem_adjm}, the above statement implies that the error in computing the $n$ Legendre coefficient of
the density for the Neumann problem, is the same as error in computing the solution of a Dirichlet problem with 
data given by a Legendre polynomial on the same edge.  This seems problematic since the right hand side
is not smooth, however, in the special case of polygonal domains, this problem is alleviated since the kernel corresponding
to the double layer potential is identically zero on line segments.

Let $V$ denote the collection of corner points in parameter space $[0,L]$, i.e. $a \in V$ if $\gamma(a)$ is a corner vertex.
Recall that $\sigma_{f}$ denote the solution to the Dirichlet problem with boundary data $f$, i.e., 
$\sigma_{f}$ satisfies
\begin{equation}
\sigma_{f}(s)   + \int_{0}^{L} k(s,t) \sigma_{f}(t) dt  = f(s) \,  \quad s\in[0,L]\setminus{V}.
\end{equation}
 Then $\sigma_{f} = f + \tilde{\sigma}$ where $\tilde{\sigma}$ satisfies the integral equation
 \begin{equation}
-\tilde{\sigma}+ \int_{0}^{L} k(s,t) \tilde{\sigma}(t) dt = \int_{s_{1}}^{s_{2}} k(s,t) f(t) dt \, , \quad s\in[0,L] \setminus V \, ,
 \end{equation}
i.e, $\tilde{\sigma}$ is the solution of Dirichle with data at $\tilde{f}$ given by
\begin{equation}
\tilde{f}(s) = \int_{s_{1}}^{s_{2}} k(s,t) f(t) dt \, , s \in [0,L] \setminus V \, .
\end{equation}
Now there are two concerns that need to be addressed. First, the accuracy of computing $\tilde{f}(s)$ 
for any point $s \in [0,L]$ using an $M$ point Gauss-Legendre quadrature on $[s_{1},s_{2}]$, and secondly, 
the resolution of the function $\tilde{f}(s)$ on the given discretization of the boundary.

For any $s$ which is contained on the same segment as $\gamma_{[s_{1},s_{2}]}$, the kernel $k(s,t)$ is identically $0$.
Thus the boundary data $\tilde{f}(s)=0$  on the same edge as the panel $\gamma([s_{1},s_{2}])$. 
We further observe that $f(t)$ is an entire function when extended to the complex plane (since it is a Legendre polynomial)
and the nearest singularity of the function $k(s,t)$ in the complex plane as a function of $t$ is corresponding to $\gamma(s)$.
Since we have assumed that no other point on the boundary $\Gamma$ is contained in the $\rho-$ bernstein ellipse
of the panel $\gamma[(s_{1},s_{2})]$, 
the error in computing $\tilde{f}(s)$ using an $M$ point Gauss-Legendre rule is bounded by $C \rho^{-M}$, where
the constant $C$ is related to the boundedness of the function $k(s,t)$ when $t$ is complexified.
(Note the error is not $\rho^{-2M}$ since $P_{M}$ would grow like $\rho^{M}$ on the $\rho-$ Bernstein ellipse.) 
Thus the function $\tilde{f}(s)$ can be computed to any desired precision by increasing the order of quadrature nodes
used to compute the integrals.

With regards to the resolution of the of the function $\tilde{f}(s)$ on the given discretization of the boundary, 
we note that the closest singularity of the function $\tilde{f}(s)$ when restricted to a panel away from the corner
and not on the same edge as $\gamma([s_{1},s_{2}])$ is the closest point on the panel $\gamma([s_{1},s_{2}])$. 
However, since we have assumed that all panels away from corners are such that no point on $\gamma([s_{1},s_{2}])$ 
is contained in their $\rho-$ Bernstein ellipse, the error in resolving the function $\tilde{f}(s)$ using an $M$ point Gauss-Legendre
or Chebyshev panel is bounded by $C \rho^{-M}$. Note that the behavior of $\tilde{f}(s)$ in the complex plane is related
to the behavior of $k(s,t)$ in the complex $s$ plane and hence can be made $O(1)$.
For the panels, at the corner, based on the proof in Appendix A, the error in resolving the function $\tilde{f}(s)$ when
truncated to a taylor series of order $N$ is less than $C 2^{-N}$, since all points on $\gamma([s_{1},s_{2}])$ are 
well-separated from corners by twice the panel length $r_{c}$. Thus, by making the panels small enough, $\rho$ can be increased arbitrarily to obtain desired tolerances on the boundary data $\tilde{f}(s)$ on the corresponding discretization of the boundary.

 Thus, the boundary data $\tilde{f}(s)$ is piecewise analytic , which can be approximated to any desired tolerance by appropriately reducing the panel sizes. This is the precise setup for which the discretization of the Dirichlet problem is designed to obtain accurate solutions to the density $\tilde{\sigma}$. 
\end{document}
